Tampering Detection In Low-Power Smart Cameras

### 2015-05-28 paper.tex
tolti i related

\begin{abstract}
A desirable feature for smart cameras is the ability to autonomously detect any tampering event/attack that would prevent a clear view over the monitored scene. No matter whether tampering is due to atmospheric phenomena (e.g., few rain drops over the camera lens) or to malicious attacks (e.g., the device displacement), these have to be promptly detected to possibly activate countermeasures. Tampering detection becomes particularly challenging in battery-powered cameras, where it is not possible to acquire images at video-like frame-rates, nor use sophisticated image-analysis algorithms.

We here introduce a tampering-detection algorithm specifically designed for low-power smart cameras: the algorithm leverages very simple indicators that are then monitored by an outlier-detection scheme. Any frame yielding anomalous indicator is detected as a tampering attempt. Core of the algorithm is the partitioning of the scene into adaptively defined regions, that are preliminarily defined by segmenting the image during the algorithm-configuration phase, and which shows to substantially improve the detection of camera displacements. Our experiments show that the proposed algorithm can successfully operate on sequences acquired at very low-frame rate, such as one frame every minute, with a very small computational complexity. %and leveraging the image partitioning into regions yields improved performance with respect to monitoring the whole scene

\keywords{tampering detection, defocus, displacement detection}
\end{abstract}

\section{Introduction}\label{sec:introduction}
%\gi{Adriano: decidi che email tenere, non tutte e due :)}

%We here address the problem of detecting tampering events/attacks in camera devices, and in particular those preventing the proper acquisition of the monitored scene.

% What is the problem?
When cameras operate outdoor and in harsh environments, dust, rain drops or snow flakes might lie on the camera lens resulting in blurry pictures, as in Fig. \ref{fig:pioggia}, or in partial occlusions of the scene, as in Fig. \ref{fig:neve}. Similarly, intentional attacks like displacing the camera, changing its focus, or spraying some opaque or glossy liquid over the lenses, would result in heavily compromised pictures, that would be surely useless for monitoring purposes. We refer to these events/attacks as tampering. In some cases, tampering is easy to detect, e.g., when the camera integrity is affected and the device goes out-of-order. However, in many other situations, namely when the device is not physically damaged, tampering detection is not straightforward, and sometimes image analysis is the only viable option.

@@

% Why is it interesting and important?
Tampering detection is an essential feature in surveillance systems~\cite{hampapur2005smart}, where cameras are expected to autonomously detect any tampering, and promptly reporting alerts. In fact, tampering prevents the correct scene interpretation: even a mild blur might result in important details such as licence plates. Surveillance cameras are typically connected to the power supply and operates at normal frame-rates (e.g. around few frames per second): in these conditions, several tampering-detection algorithms have been presented~\cite{TODO}.

In this work we expressly target low-power and ultra-low-power smart cameras\gi{una reference qua ci vuole}, battery-powered devices that are characterized by a constrained computational power and memory availability. Even though these devices are not meant for critical surveillance applications, they are becoming popular in distributed systems for monitoring wide environments due to their low cost and maintenance requirements.  %As an example, consider Wireless Multimedia Sensor Networks (WMSN) \cite{akyildiz2007survey} where smart-cameras can acquire and transmit images at regular time interval or upon requests. 
% These units are not connected to the power supply and have to operate with batteries and possibly rely on energy harvesting~\cite{magno2009adaptive}. 
%In practice, because of energy constrain, low-power devices have to operate at very low frame-rates: possibly less than one frame every minute.

% Why is it hard? (E.g., why do naive approaches fail?)
Tampering detection in low-power smart cameras has not been much investigated in the literature~\cite{alippi2010detecting} while being more challenging than in conventional surveillance cameras~\cite{perrig2004security}. Beside computational aspects -- such as the number of operations per pixels allowed -- the big issue is that low-power smart cameras typically operate at very low-frame rates (e.g., less than one frame per minute), thus the acquired sequence does not evolve smoothly. This prevents the use of learned background models and the analysis of foreground variations \cite{piccardi2004background}. In fact, when dynamic environments are acquired at low frame rates, two consecutive frames might be very different because of changes in the scene and in the light conditions, yielding sequences like those in Figure~\ref{fig:sequences}. Smart cameras have to promptly distinguish between \emph{normal} scene or illumination changes, and changes due to camera tampering to avoid the transmission of corrupted frames. Moreover, the false-alarm rate is a serious concern, a any false alarm results in wasting energy for data processing and network transmissions.
% We here address the problem of detecting tampering events on a camera device that is employed for monitoring purpose. Tampering events can be due to malicious attacks, as for instance a displacement of the camera or spraying some dirty liquid that would prevent the acquisition of (part of) the monitored scene or the interpretation of the content (e.g., the identification of licence plates).

%What are the key components of my approach and results? Also include any specific limitations.
We here address the detection of camera blurring and displacement, which are presented in Section \ref{sec:probForm}. The proposed algorithm, detailed in Section~\ref{sec:propSol}, relies on two indicators: the average image intensity to detect displacement, and the average gradient norm, to detect blurring. These indicators can be easily computed and monitored by an outlier-detection technique in a low-power smart camera, thus our solution is meant as a very prompt trigger to be possibly combined with other sequential monitoring techniques. In particular, we show that separately monitoring the indicators over different regions of the scene can substantially improve the displacement-detection performance. Remarkably, these regions can be preliminarily defined during an initial configuration phase (Section \ref{subsec:Segmentation}), thus the algorithm operates at a negligible computational overhead w.r.t. monitoring the whole image. Experiments in Section \ref{sec:experiments} show that leveraging image regions can substantially improve the displacement-detection performance. % Concluding remarks and discussions are given in Section \ref{sec:Conclusion}.

\subsection{Related Works}\label{subsec:relWorks}
% Why hasn't it been solved before? (Or, what's wrong with previous proposed solutions? How does mine differ?)
The literature concerning tampering detection is mostly focused on video surveillance applications and operates at few frames per second \gi{Adriano, \'e vero?}. Background models are typically leveraged to identify defocus and occlusions; in particular~\cite{aksay2007camera} performs defocus detection by analyzing the wavelet domain of each frame and performs histogram comparison for detecting occlusions, while camera displacements are not considered. A background-subtraction technique is employed in~\cite{saglam2009real} to identify defocus, occlusions, and displacements. \gi{FIXME Comparison in the Fourier domain for defocus detection, histogram comparison for occlusion detection, comparison between current background and delayed background for displacement detection}. Background subtraction is also used in \cite{gil2007automatic} to detect defocus, occlusions, and displacements. \gi{FIXME Comparison of edges pixels count for defocus detection, entropy comparison for occlusion detection, block matching algorithm for displacement detection}.
In contrast, no background models are used in \cite{alippi2010detecting}, where a sequential monitoring scheme based on a change-detection test is employed to detect changes in the average gradient energy of each frame to detect defocus due to external disturbances on the camera lens.  
\cite{ribnick2006real}: comparison between frames belonging to a buffer in order to find high values of dissimilarity, associated to tampering.
\cite{kryjak2012fpga}: implementation in a FPGA of a solution based on background modeling, histograms comparisons, edges comparisons.
\cite{harasse2004automated}: tampering detection inside a moving vehicle; uses background subtraction methods in order to identify defocus, occlusions, and displacements. Comparison of edges pixels count for defocus detection, entropy comparison for occlusion detection, block matching algorithm for displacement detection.
\cite{tsesmelis2013tamper}: monitoring of the number of key points extracted by SURF in order to detect defocus events, partition in blocks and HOG descriptors matching for each block in order to detect occlusions. These types of solutions requires a lot of computations
%
%
%
\section{Problem Formulation}\label{sec:probForm}
%
Let $z_t$ be the frame acquired at time $t$
\begin{equation}
\label{eq:observationModel}
z_t(x)=\mathcal{D}_t[y_t](x) \quad \forall x \in X
\end{equation}
where $\mathcal{D}_t$ denotes an operator that transforms the original image $y_t$ in the acquired frame $z_t$; $X \subset \mathbb{Z}^2$ denotes the regular pixels grid and $x\in \mathbb{Z}^2$ indicates the pixel coordinates. As far as there is no tampering attacks/events,
\begin{equation}
\label{eq:no_tampering}
\mathcal{D}_t[y_t](x) = y_t(x) + \eta_t(x) \quad \forall x \in X
\end{equation}
where $\eta_t$ is a random variable accounting for image noise, %sources (e.g., thermal, quantization, photon-counting). 
and all the images $y_t$ might show different content but are acquired from the same viewpoint and the same camera orientation.

@@

We have acquired two datasets to test our method. The first dataset contains \gi{how many?} sequences taken from webcams recording urban areas (as in Figure \ref{fig:sequences}(a)), where we have synthetically introduced blurring by means of a gaussian filters (Figure \ref{fig:sequences}(b)), and displacement by concatenating frames from sequences depicting similar environments (Figure \ref{fig:sequences}(c)). In some cases, sequences contained real tampering events, as in Figure \ref{fig:sequences}(d).\\
\begin{figure}[t]
\centering
\includegraphics[width=1\linewidth]{Immagini/sequenze}
\caption{Sequences taken from webcams: \textbf{(a)} no tampering events; \textbf{(b)} defocus event on 4-th and 5-th frames, created using gaussian filtering; \textbf{(c)} displacement event on 4-th and 5-th frames, created using concatenation between similar sequences; \textbf{(d)} real displacement event on 4-th frame.}
\label{fig:sequences}
\end{figure}
The second dataset was recorded using a Raspberry Pi Model B+, with its camera module, and the tampering was introduced by moving the device or spraying water on the camera len.\gi{una footnote dicendo sequences can be released upon request?}

@@

\section{Conclusion}\label{sec:Conclusion}
\gi{Adriano: butta in inglese gli ongoing works (come ultima cosa)}
.

>>>>

\begin{abstract}
A desirable feature in smart cameras is the ability to autonomously detect any tampering event/attack that would prevent a clear view over the monitored scene. No matter whether tampering is due to atmospheric phenomena (e.g., few rain drops over the camera lens) or to malicious attacks (e.g., the device displacement), these have to be promptly detected to possibly activate countermeasures. Tampering detection becomes particularly challenging in battery-powered cameras, where it is not possible to acquire images at video-like frame-rates, nor use sophisticated image-analysis algorithms.

We here introduce a tampering-detection algorithm specifically designed for low-power smart cameras: the algorithm leverages very simple indicators that are then monitored by an outlier-detection scheme. Any frame yielding anomalous indicator is detected as a tampering attempt. Core of the algorithm is the partitioning of the scene into adaptively defined regions, that are preliminarily defined by segmenting the image during the algorithm-configuration phase, and which shows to substantially improve the detection of camera displacements. Our experiments show that the proposed algorithm can successfully operate on sequences acquired at very low-frame rate, such as one frame every minute, with a very small computational complexity. %and leveraging the image partitioning into regions yields improved performance with respect to monitoring the whole scene
\keywords{tampering detection, blurring detection, displacement detection}
\end{abstract}

\section{Introduction}\label{sec:introduction}
% What is the problem?
When cameras operate outdoor and in harsh environments, dust, rain drops or snow flakes might lie on the camera lens resulting in blurry pictures, as in Fig. \ref{fig:pioggia}, or in partial occlusions of the scene, as in Fig. \ref{fig:neve}. Similarly, intentional attacks like displacing the camera, changing its focus, or spraying some opaque or glossy liquid over the lenses, would result in heavily compromised pictures, that would be useless for monitoring purposes. We refer to these events/attacks as tampering. In some cases, tampering is easy to detect, e.g., when the camera integrity is affected and the device goes out-of-order. However, in many other situations, namely when the device is not physically damaged, tampering detection is not straightforward, and sometimes image analysis is the only viable option.

@@

% Why is it interesting and important?
Tampering detection is an essential feature in surveillance systems~\cite{hampapur2005smart}, where cameras are expected to autonomously detect any tampering, and promptly report alerts. In fact, tampering prevents the correct scene interpretation: even a mild blur might hinder the identification of important details such as licence plates. Surveillance cameras are typically connected to the power supply and operates at normal frame-rates (e.g. around few frames per second): in these conditions, several tampering-detection algorithms have been presented.

In particular, camera displacement and occlusions are typically detected by background models or frame comparison, while blurring is instead detected by monitoring the high-frequency component of images, as this is expected to drop. In ~\cite{aksay2007camera} and~\cite{saglam2009real} tampering is detected by comparing histograms of background and current frame, and by monitoring the energy in wavelet or Fourier domain. Two background models computed over different time intervals are used in~\cite{saglam2009real}. In~\cite{gil2007automatic}, tampering detection is performed by combining background subtraction together with edge detection and normalized cross-correlation. Background models and block matching in~\cite{tsesmelis2013tamper} enable the displacement detection, while number of SURF~\cite{SURF} keypoints is used to detect blurring. A frame buffer rather than an explicit background is used~\cite{ribnick2006real}. All these solutions, including~\cite{ribnick2006real,harasse2004automated,kryjak2012fpga} are computationally demanding, and are not viable options for low-power cameras.

In this work we expressly target low-power and ultra-low-power smart cameras\gi{Claudio, riesci a mettere una reference qua?}, battery-powered devices that are characterized by a constrained computational power and memory availability. Even though these devices are not meant for critical surveillance applications, they are becoming popular in distributed systems for monitoring wide environments due to their low cost and maintenance requirements.  %As an example, consider Wireless Multimedia Sensor Networks (WMSN) \cite{akyildiz2007survey} where smart-cameras can acquire and transmit images at regular time interval or upon requests. 
% These units are not connected to the power supply and have to operate with batteries and possibly rely on energy harvesting~\cite{magno2009adaptive}. 
%In practice, because of energy constrain, low-power devices have to operate at very low frame-rates: possibly less than one frame every minute.

% Why is it hard? (E.g., why do naive approaches fail?)
Tampering detection in low-power smart cameras has not been much investigated in the literature~\cite{alippi2010detecting}, while being more challenging than in conventional surveillance cameras~\cite{perrig2004security}. Beside computational aspects -- such as the number of operations per pixels allowed -- the big issue is that low-power smart cameras typically operate at very low-frame rates (e.g., less than one frame per minute), thus the acquired sequence does not evolve smoothly. This prevents the use of learned background models and the analysis of foreground variations \cite{piccardi2004background}. In fact, when dynamic environments are acquired at low frame rates, two consecutive frames might be very different because of changes in the scene and in the light conditions, yielding sequences like those in Figure~\ref{fig:sequences}. Smart cameras have to promptly distinguish between \emph{normal} scene or illumination changes, and changes due to camera tampering to avoid the transmission of corrupted frames. Moreover, false alarms represent a serious problem in low-power smart cameras as these result in useless data processing and network transmissions, thus energy waste.
% We here address the problem of detecting tampering events on a camera device that is employed for monitoring purpose. Tampering events can be due to malicious attacks, as for instance a displacement of the camera or spraying some dirty liquid that would prevent the acquisition of (part of) the monitored scene or the interpretation of the content (e.g., the identification of licence plates).

%What are the key components of my approach and results? Also include any specific limitations.
We address the detection of camera blurring and displacement, which are presented in Section \ref{sec:probForm}. The proposed algorithm, detailed in Section~\ref{sec:propSol}, relies on two indicators: the average image intensity is used to detect displacement, and the average gradient norm to detect blurring. These indicators can be easily computed and monitored by an outlier-detection technique in a low-power smart camera. In particular, we show that separately monitoring the indicators over different regions of the image can substantially improve the displacement-detection performance. Remarkably, these regions can be preliminarily defined during an initial configuration phase (Section \ref{subsec:Segmentation}), thus the algorithm operates at a negligible computational overhead w.r.t. monitoring the whole image. Thus, our solution represents as a prompt trigger that can be possibly combined with other sequential monitoring techniques. Experiments in Section \ref{sec:experiments} show that leveraging image regions can substantially improve the displacement-detection performance. % Concluding remarks and discussions are given in Section \ref{sec:Conclusion}.

\begin{figure}[t]
\centering
\includegraphics[width=1\linewidth]{Immagini/sequenze}
\caption{Sequences taken from webcams: \textbf{(a)} no tampering events; \textbf{(b)} defocus event on 4-th and 5-th frames, created using gaussian filtering; \textbf{(c)} displacement event on 4-th and 5-th frames, created using concatenation between similar sequences; \textbf{(d)} real displacement event on 4-th frame.}
\label{fig:sequences}
\end{figure}

%A low-complexity and sequential monitoring scheme over the average gradient norm is used in~\cite{alippi2010detecting} to detect blurring due to external disturbances on the camera lens.

%\subsection{Related Works}\label{subsec:relWorks}
%% The literature concerning tampering detection is mostly focused on video surveillance applications and process at few frames per second.
% 
%%
%: comparison between frames belonging to a buffer in order to find high values of dissimilarity, associated to tampering.
%\cite{harasse2004automated}: tampering detection inside a moving vehicle; uses background subtraction methods in order to identify defocus, occlusions, and displacements. Comparison of edges pixels count for defocus detection, entropy comparison for occlusion detection, block matching algorithm for displacement detection.
%
%
%
\section{Problem Formulation}\label{sec:probForm}
%
Let $z_t$ be the frame acquired at time $t$
\begin{equation}
\label{eq:observationModel}
z_t(x)=\mathcal{D}_t[y_t](x) \quad \forall x \in X
\end{equation}
where $\mathcal{D}_t$ denotes an operator that transforms the original image $y_t$ in the acquired frame $z_t$; $X \subset \mathbb{Z}^2$ denotes the regular pixels grid and $x\in \mathbb{Z}^2$ indicates the pixel coordinates. As far as there is no tampering attacks/events,
\begin{equation}
\label{eq:no_tampering}
\mathcal{D}_t[y_t](x) = y_t(x) + \eta_t(x) \quad \forall x \in X
\end{equation}
where $\eta_t$ is a random variable accounting for image noise, %sources (e.g., thermal, quantization, photon-counting). 
and all the images $y_t$ might show different content but are acquired from the same viewpoint and the same camera orientation.

@@

We have acquired two datasets to test our method. The first dataset contains \gi{how many?} sequences taken from webcams recording urban areas (as in Figure \ref{fig:sequences}(a)), where we have synthetically introduced blurring by means of a gaussian filters (Figure \ref{fig:sequences}(b)), and displacement by concatenating frames from sequences depicting similar environments (Figure \ref{fig:sequences}(c)). In some cases, sequences contained real tampering events, as in Figure \ref{fig:sequences}(d).\\

The second dataset was recorded using a Raspberry Pi Model B+, with its camera module, and the tampering was introduced by moving the device or spraying water on the camera len.\gi{una footnote dicendo sequences can be released upon request?}

@@

\section{Conclusion}\label{sec:Conclusion}
A low-complexity and sequential monitoring scheme over the average gradient norm is used in~\cite{alippi2010detecting} to detect blurring due to external disturbances on the camera lens.
